---
title: "Predicting H1N1 vaccine likelihood using Data Mining Methods"
author: "Luke Awino, Roberto Cancel, & Kevin Stewart"
date: "7/27/2021"
output:
  word_document: default
  html_document:
    df_print: paged
---

**Team: 6**  
**Data set: "Flu Shot Learning: "Predict H1N1 and Seasonal Flu Vaccines"**  
**Origin: "UCI Machine Learning Repository"**  
**Objective: The goal is to predict the probability of individuals getting their H1N1 vaccine using behavioral and demographic information.**  
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(car)
library(carData)
library(corrplot)
library(caret)
library(Hmisc)
library(plyr)
library(dplyr)
library(lattice)
library(ggplot2)
library(mice)
library(naniar)
library(simputation)
library(visdat)
library(MASS)
library(mlbench)
library(performance)
library(reshape2)
library(pROC)
library(rpart)
library(naivebayes)
library(e1071)
library(kernlab)
library(randomForest)
memory.limit(1000000)
```
 
# Data Importing and Pre-processing
 
*Import the Training data set*
```{r message=FALSE, include=FALSE}
#Import the feature data set
h1n1_df <- read.csv('training_set_features.csv', header=TRUE, row.names="respondent_id", na.strings=c(""," ","NA"))
#import response/target variables
targ_var <- read.csv("training_set_labels.csv", header=TRUE, row.names="respondent_id")
#Add the target variable to our data set
h1n1_df$h1n1_vaccine <- targ_var$h1n1_vaccine
head(h1n1_df)
```
*Examine the structure of the data set*
```{r}
#Look at the the structure of the data
str(h1n1_df)
```
*Examine missing values for first round of feature elimination*
```{r}
# sort missing values by count 
describe(h1n1_df)
```
*Remove features with large proportion of missing data*
```{r}
#Removing employment data (since 13330/26707 or 50% of employment_industry is missing and 13470/26707 or 50% of employment_occupation is missing) and health_insurance (50% missing) and hhs_geo_region to focus on Census_msa
h1n1_df <- subset(h1n1_df, select = -c(hhs_geo_region, employment_industry, employment_occupation, health_insurance))
```
*Review Missing Data still in df*
```{r}
# Count missing data in the data frame
sort(colSums(is.na(h1n1_df)))
```
*Impute Missing Values for Categorical Variables with mode*
```{r}
h1n1_df <- h1n1_df[complete.cases(h1n1_df), ]
str(h1n1_df)
```
```{r}
#Verify that all the data is is not missing 
sort(colSums(is.na(h1n1_df)))
```
*Transform the features*
```{r}
#converting categorical variables to factors 
h1n1_df$education <- as.factor(h1n1_df$education)
h1n1_df$race <- as.factor(h1n1_df$race)
h1n1_df$sex <- as.factor(h1n1_df$sex)
h1n1_df$age_group <- as.factor(h1n1_df$age_group)
h1n1_df$income_poverty <- as.factor(h1n1_df$income_poverty)
h1n1_df$marital_status <- as.factor(h1n1_df$marital_status)
h1n1_df$rent_or_own <- as.factor(h1n1_df$rent_or_own)
h1n1_df$employment_status <- as.factor(h1n1_df$employment_status)
#converting integers discrete variables to factors
h1n1_df$h1n1_concern <- as.factor(h1n1_df$h1n1_concern)
h1n1_df$h1n1_knowledge <- as.factor(h1n1_df$h1n1_knowledge)
h1n1_df$behavioral_antiviral_meds <- as.factor(h1n1_df$behavioral_antiviral_meds)
h1n1_df$behavioral_avoidance <- as.factor(h1n1_df$behavioral_avoidance)
h1n1_df$behavioral_face_mask <- as.factor(h1n1_df$behavioral_face_mask)
h1n1_df$behavioral_wash_hands <- as.factor(h1n1_df$behavioral_wash_hands)
h1n1_df$behavioral_large_gatherings <- as.factor(h1n1_df$behavioral_large_gatherings)
h1n1_df$behavioral_outside_home <- as.factor(h1n1_df$behavioral_outside_home)
h1n1_df$behavioral_outside_home <- as.factor(h1n1_df$behavioral_touch_face)
h1n1_df$behavioral_touch_face <- as.factor(h1n1_df$behavioral_touch_face)
h1n1_df$doctor_recc_h1n1 <- as.factor(h1n1_df$doctor_recc_h1n1)
h1n1_df$doctor_recc_seasonal <- as.factor(h1n1_df$doctor_recc_seasonal)
h1n1_df$chronic_med_condition <- as.factor(h1n1_df$chronic_med_condition)
h1n1_df$child_under_6_months <- as.factor(h1n1_df$child_under_6_months)
h1n1_df$health_worker <- as.factor(h1n1_df$health_worker)
h1n1_df$opinion_h1n1_vacc_effective <- as.factor(h1n1_df$opinion_h1n1_vacc_effective)
h1n1_df$opinion_h1n1_risk <- as.factor(h1n1_df$opinion_h1n1_risk)
h1n1_df$opinion_h1n1_sick_from_vacc <- as.factor(h1n1_df$opinion_h1n1_sick_from_vacc)
h1n1_df$opinion_seas_vacc_effective <- as.factor(h1n1_df$opinion_seas_vacc_effective)
h1n1_df$opinion_seas_risk <- as.factor(h1n1_df$opinion_seas_risk)
h1n1_df$opinion_seas_sick_from_vacc <- as.factor(h1n1_df$opinion_seas_sick_from_vacc)
h1n1_df$household_adults <- as.factor(h1n1_df$household_adults)
h1n1_df$household_children <- as.factor(h1n1_df$household_children)
h1n1_df$census_msa <- as.factor(h1n1_df$census_msa)
```

```{r}
clean_data <- h1n1_df
str(clean_data)
```
 
#Visualize categorical variables
 
```{r}
#graph 
ggplot(clean_data, aes(marital_status)) + geom_bar(aes(fill = marital_status)) + coord_flip() + ggtitle("Frequency of Marital Status")

```
```{r}
ggplot(clean_data, aes(rent_or_own)) + geom_bar(aes(fill = rent_or_own)) + coord_flip()+ ggtitle("Frequency of Rent or Own")
```
```{r}
ggplot(clean_data, aes(age_group)) + geom_bar(aes(fill = age_group)) + coord_flip() + ggtitle("Frequency of Age Group")
```
```{r}
ggplot(clean_data, aes(education)) + geom_bar(aes(fill = education)) + coord_flip()+ ggtitle("Frequency of Education Level")
```
```{r}
ggplot(clean_data, aes(race)) + geom_bar(aes(fill = race)) + coord_flip()+ ggtitle("Frequency of Race")
```
```{r}
ggplot(clean_data, aes(sex)) + geom_bar(aes(fill = sex)) + coord_flip()+ ggtitle("Frequency of Sex")
```

```{r}
ggplot(clean_data, aes(income_poverty)) + geom_bar(aes(fill = income_poverty)) + coord_flip() + ggtitle("Frequency of Income Type")
```
```{r}
ggplot(clean_data, aes(census_msa)) + geom_bar(aes(fill = census_msa)) + coord_flip() + ggtitle("Frequency of MSA type")
```
# Visualize the categorical variables as functions of h1n1_vaccine
```{r}
library(forcats)
ggplot(clean_data, aes(fct_infreq(marital_status))) +
geom_bar(stat="count", aes(fill= h1n1_vaccine)) +
labs(x = "Marital Status", y = "Count") +
ggtitle("Marital Status by Performance: (Vaccinated or Not Vaccinated)")
```

**FEATURE ENGINEERING**
 *Re-express Categorical Variables and convert to numerical*
```{r}
unique(clean_data$marital_status)
unique(clean_data$income_poverty)
unique(clean_data$rent_or_own)
unique(clean_data$education)
unique(clean_data$employment_status)
```
```{r}
# Re-expressing categorical variables as a value 
marital_num <- revalue(x = clean_data$marital_status, replace = c("Not Married" = 0, "Married" = 1))
clean_data$marital_numeric <- as.numeric(levels(marital_num))[marital_num]

# Re-expressing census_msa
census_msa <- as.factor(clean_data)
census_msa_num <- census_msa_num <- revalue(x = clean_data$census_msa, replace = c("Non-MSA" = 0, "MSA, Not Principle  City" = 1, "MSA, Principle City" = 2))
clean_data$census_msa_numeric <- as.numeric(levels(census_msa_num))[census_msa_num]

# Re-expressing age as numeric
unique(clean_data$age_group)
length(unique(clean_data$age_group))
age_num <- revalue(x = clean_data$age_group, replace = c("18 - 34 Years" = 0, "35 - 44 Years" = 1, "45 - 54 Years" = 2, "55 - 64 Years" = 3, "65+ Years" = 4))
# convert age_num to numeric
clean_data$age_numeric <- as.numeric(levels(age_num))[age_num]

#Re-express sex as numeric
sex_num <- revalue(x = clean_data$sex, replace = c("Female" = 0, "Male" = 1))
clean_data$sex_numeric <- as.numeric(levels(sex_num))[sex_num]

#convert race to numeric
unique(clean_data$race)
race_num <- revalue(x = clean_data$race, replace = c("White" = 0, "Black" = 1, "Other or Multiple" = 2, "Hispanic" = 3))
clean_data$race_numeric <- as.numeric(levels(race_num))[race_num]

#converting income_poverty to numeric
income_poverty_num <- revalue(x = clean_data$income_poverty, replace = c("Below Poverty" = 0, "<= $75,000, Above Poverty" = 1, "> $75,000" = 2))
clean_data$income_poverty_numeric <- as.numeric(levels(income_poverty_num))[income_poverty_num]

#Re-expressing categorical variables
unique(clean_data$rent_or_own)
rent_or_own_num <- revalue(x = clean_data$rent_or_own, replace = c("Own" = 0, "Rent" = 1))
clean_data$rent_or_own_numeric <- as.numeric(levels(rent_or_own_num))[rent_or_own_num]

#Re-expressing categorical variables 
unique(clean_data$education)
education_num <- revalue(x = clean_data$education, replace = c("< 12 Years" = 0, "12 Years"= 1, "College Graduate" = 2, "Some College"= 3))
clean_data$education_numeric <- as.numeric(levels(education_num)) [education_num]

#Re-expressing categorical variables 
unique(clean_data$employment_status)
employment_num <- revalue(x = clean_data$employment_status, replace = c("Unemployed" = 0, "Not in Labor Force" = 1, "Employed" = 2))
clean_data$employment_numeric <- as.numeric(levels(employment_num)) [employment_num]


#Re-express categorical variables 
str(clean_data)
```
*Drop the Categorical Variables and Seasonal Flu data since we're focusing on H1N1 vaccines*
```{r}
clean_data1 <- subset(clean_data, select = -c( age_group, education, race, sex, income_poverty, marital_status, rent_or_own, employment_status, census_msa, doctor_recc_seasonal, opinion_seas_vacc_effective, opinion_seas_risk, opinion_seas_sick_from_vacc))
```
*Convert all variables to numeric after transformations*
```{r}
prep_data <- mutate_all(clean_data1, function(clean_data)as.numeric(clean_data))
str(prep_data)
```
```{r}
#View updated prep_data
str(prep_data)
```
 
#Final preparation before modeling
 
 *Check for multicollinearity in features*
```{r}
# calculate correlation matrix
cormat <- round(cor(prep_data, method = "spearman"), 2)
# Melt the cormat
melted_cormat <- melt(cormat)
# Get lower triangle of the correlation matrix
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
#Get upper tri
upper_tri <- get_upper_tri(cormat)

#Create clearer correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)

#Create heat map
ggheatmap <- ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Spearman\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 90, vjust = 1, 
    size =6, hjust = 1))+
 coord_fixed()

#Add Coefficients
ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 2) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
```                
*Scale all the features*
```{r}
#Scale ordinal features
max = apply(prep_data,2, max)
min = apply(prep_data,2, min)
prep_data = as.data.frame(scale(prep_data, center = min, scale = max - min))
```

*Partition the data*
```{r}
#Set Seed and determine dimensions of data set
set.seed(654)
n <- dim(prep_data)
n
```
```{r}
# Split the data into 75% train and 25% test
dt = sort(sample(nrow(prep_data), nrow(prep_data)*.75))

prep_train <- prep_data[dt,]
prep_test <- prep_data[-dt,]
```

*Balance the training set*

```{r}
#Count number of records in train set
dim(prep_train)
#Count number of records in test set
dim(prep_test)
```
*Identify number of h1n1_vaccine is True in training set*

```{r}
length(which(prep_train$h1n1_vaccine == "1"))
```
 
There are 20030 records in the training data set, of which 4273 have a h1n1_vaccine of True/1 - this means only 21% of the training set has h1n1_vaccine of True/1. We would like to balance the training set to a 50/50 of h1n1_vaccine True/1 and False/0.

To reach this x = ((.5*14731)-3409)/.5
x = 7913

We therefore need to over sample the h1n1_vaccine True/1 records by 11544 to balance our data set.
 

*Balance the training data set for the imbalance in H1N1_vaccine*
```{r}
# Define the records to be sample from
to.resample <- which(prep_train$h1n1_vaccine == "1")
# Build a sample of size 11,544 from identified records
our.resample <- sample(x = to.resample, size = 7913, replace = TRUE)
our.resample <- prep_train[our.resample,]
# Bind re-sampled records with training data
prep_train_rebal <- rbind(prep_train, our.resample)
# Build Table of Response Counts and Proportions
t.v1 <- table(prep_train_rebal$h1n1_vaccine)
t.v2 <- rbind(t.v1, round(prop.table(t.v1), 2))
colnames(t.v2) <- c("h1n1_vaccine = False/0", "h1n1_vaccine = True/1")
rownames(t.v2) <- c("Count", "Proportion")
t.v2
```

```{r}
str(prep_train_rebal)
```
 
#Modeling#
*Logistic Regression*
```{r}
#Build and train baseline model with all remaining features

logreg01 <- glm(formula = h1n1_vaccine ~ h1n1_concern + h1n1_knowledge + behavioral_antiviral_meds + behavioral_avoidance + behavioral_face_mask + behavioral_wash_hands + behavioral_large_gatherings + behavioral_outside_home + doctor_recc_h1n1 + chronic_med_condition + child_under_6_months + health_worker + opinion_h1n1_vacc_effective + opinion_h1n1_risk + opinion_h1n1_sick_from_vacc + household_adults + household_children + marital_numeric + census_msa_numeric + age_numeric + sex_numeric + race_numeric + income_poverty_numeric + rent_or_own_numeric + education_numeric + employment_numeric, 
    data = prep_train_rebal, family = binomial(link = "logit"))

summary(logreg01)
```
 
Summary: We see that the following variables are statistically insignificant and therefore, likely, do not significantly contribute to the likelihood of vaccination:  behavioral_outside_home, opinion_h1n1_sick_from_vacc, household_adults, census_msa_numeric, race_numeric, rent_or_own_numeric, education_numeric, employment_numeric. We have, however, decided to keep race, education and employment in our subsequent iteration since this is a socio-demographic study.
 
*Validate with the test set*
```{r}
logreg01_test <- glm(formula = h1n1_vaccine ~ h1n1_concern + h1n1_knowledge + behavioral_antiviral_meds + behavioral_avoidance + behavioral_face_mask + behavioral_wash_hands + behavioral_large_gatherings + behavioral_outside_home + doctor_recc_h1n1 + chronic_med_condition + child_under_6_months + health_worker + opinion_h1n1_vacc_effective + opinion_h1n1_risk + opinion_h1n1_sick_from_vacc + household_adults + household_children + marital_numeric + census_msa_numeric + age_numeric + sex_numeric + race_numeric + income_poverty_numeric + rent_or_own_numeric + education_numeric + employment_numeric, 
    data = prep_test, family = binomial(link = "logit"))

summary(logreg01_test)
```

*Obtain the predicted values of the target variable for each record in the data set
```{r}
pred = predict(logreg01, newdata=prep_test)
predicted.classes <- factor(ifelse(pred > 0.5, "1", "0"))
accuracy <- table(pred, prep_test[,"h1n1_vaccine"])
sum(diag(accuracy))/sum(accuracy)
```
```{r}
confusionMatrix(predicted.classes, factor(prep_test$h1n1_vaccine), positive = '1')
```

```{r}
auc(prep_test$h1n1_vaccine, pred)
```
*Rationalized Logistic Regression*
```{r}
#Build and train baseline model with all remaining features
logreg02 <- glm(formula = h1n1_vaccine ~ h1n1_concern + h1n1_knowledge + behavioral_antiviral_meds + behavioral_avoidance + behavioral_face_mask + behavioral_wash_hands + behavioral_large_gatherings + doctor_recc_h1n1 + chronic_med_condition + child_under_6_months + health_worker + opinion_h1n1_vacc_effective + opinion_h1n1_risk + household_children + marital_numeric + census_msa_numeric + age_numeric + sex_numeric + race_numeric + income_poverty_numeric + education_numeric + employment_numeric, 
    data = prep_train_rebal, family = binomial(link = "logit"))

summary(logreg02)
```

*Validate with the test set*
```{r}
logreg02_test <- glm(formula = h1n1_vaccine ~ h1n1_concern + h1n1_knowledge + behavioral_antiviral_meds + behavioral_avoidance + behavioral_face_mask + behavioral_wash_hands + behavioral_large_gatherings + doctor_recc_h1n1 + chronic_med_condition + child_under_6_months + health_worker + opinion_h1n1_vacc_effective + opinion_h1n1_risk + household_children + marital_numeric + census_msa_numeric + age_numeric + sex_numeric + race_numeric + income_poverty_numeric + education_numeric + employment_numeric,
    data = prep_test, family = binomial(link = "logit"))

summary(logreg02_test)
```
*Obtain the predicted values of the target variable for each record in the data set
```{r}
pred3 = predict(logreg02, newdata=prep_test)
predicted.classes3 <- factor(ifelse(pred3 > 0.5, "1", "0"))
accuracy3 <- table(pred3, prep_test[,"h1n1_vaccine"])
sum(diag(accuracy))/sum(accuracy)
```
```{r}
auc(prep_test$h1n1_vaccine, pred3)
```
```{r}

confusionMatrix(predicted.classes3, factor(prep_test$h1n1_vaccine), positive = '1')
```

# Naives Bayes 

```{r}
#Building a naive bayes model 

nb1 <- naiveBayes(formula = h1n1_vaccine ~ + h1n1_knowledge + 
    behavioral_antiviral_meds + behavioral_avoidance + behavioral_face_mask + 
    behavioral_wash_hands + behavioral_large_gatherings + behavioral_outside_home + 
    behavioral_touch_face + doctor_recc_h1n1 + chronic_med_condition + 
    child_under_6_months + health_worker + opinion_h1n1_vacc_effective + 
    opinion_h1n1_risk + opinion_h1n1_sick_from_vacc + household_adults + 
    household_children + marital_numeric + census_msa_numeric + 
    age_numeric + sex_numeric + race_numeric + income_poverty_numeric + 
    rent_or_own_numeric + education_numeric + employment_numeric, data = prep_train_rebal)

nb1
```
# Predictions 
```{r include=FALSE}
#Do prediction get all columns except target variable

head(prep_train_rebal)

ypred1 <- predict(nb1, newdata = prep_train_rebal)
(cbind(ypred1, prep_train_rebal))

ypred2 <- predict(nb1, newdata = prep_test)
(cbind(ypred2, prep_test))


```

# Confusion Matrix

```{r}
#Create a confusion matrix to evaluate the model.

# Confusion matrix of training set 
t.pred1 <-  table(prep_train_rebal$h1n1_vaccine, ypred1)
rownames(t.pred1) <- c("Actual:Not Vaccinated", "Actual: Vaccinated")
colnames(t.pred1) <- c("Predicted: No Vaccine", "Predicted: Vaccinated")
addmargins(A = t.pred1, FUN = list(Total=sum), quiet = TRUE)

# Confusion matrix for testing data set 
t.pred2 <-  table(prep_test$h1n1_vaccine, ypred2)
rownames(t.pred2) <- c("Actual:Not Vaccinated", "Actual: Vaccinated")
colnames(t.pred2) <- c("Predicted: Not Vaccinated", "Predicted: Vaccinated")
addmargins(A =t.pred2, FUN = list(Total=sum), quiet = TRUE)

```

# Evaluate the training model 

```{r}
require(caret)

# Convert the data to factor to run evaluations
prep_train_rebal$h1n1_vaccine <- as.factor(prep_train_rebal$h1n1_vaccine)
# verifying the data type
str(prep_train_rebal$h1n1_vaccine)
#Evaluating the confusion matrix of the training set 
pred_p <- predict(nb1, newdata=prep_train_rebal)
predicted.class_2 <- factor(ifelse(pred_p > .5, "1","0"))
accuracy1 <- table(pred_p, prep_train_rebal[,"h1n1_vaccine"])
sum(diag(accuracy1))/sum(accuracy1)
#printing the results
confusionMatrix(data=pred_p, factor(prep_train_rebal$h1n1_vaccine), positive = '1')

#Evaluating the confusion matrix of the testing data set 
pred_p <- predict(nb1, newdata=prep_test)
accuracy1 <- table(pred_p, prep_test[,"h1n1_vaccine"])
sum(diag(accuracy1))/sum(accuracy1)

#Printing the results
confusionMatrix(data=pred_p, factor(prep_test$h1n1_vaccine), positive = '1')

```


# Improving the model  by smoothing with laplace and usekernals
```{r}
str(prep_train_rebal$h1n1_vaccine)


nb2 <- naiveBayes(formula = h1n1_vaccine ~ +h1n1_knowledge + 
    behavioral_antiviral_meds + behavioral_avoidance + behavioral_face_mask +  behavioral_wash_hands + behavioral_large_gatherings +behavioral_touch_face +doctor_recc_h1n1 + chronic_med_condition + 
    child_under_6_months + health_worker + opinion_h1n1_vacc_effective + 
    opinion_h1n1_risk + opinion_h1n1_sick_from_vacc + household_adults + 
    household_children + marital_numeric + census_msa_numeric + 
    age_numeric + sex_numeric + race_numeric + income_poverty_numeric + 
    rent_or_own_numeric + education_numeric + employment_numeric, data = prep_train_rebal, laplace = 1, usekernals = 1)

nb2


```

# Predictions using smoothing and uskernals

```{r}
ypred1 <- predict(nb2, newdata = prep_train_rebal)
(cbind(ypred1, prep_train_rebal))

ypred2 <- predict(nb2, newdata = prep_test)
(cbind(ypred2, prep_test))



```


```{r}
# Confusion matrix of training set 
t.pred1 <-  table(prep_train_rebal$h1n1_vaccine, ypred1)
rownames(t.pred1) <- c("Actual:Not Vaccinated", "Actual: Vaccinated")
colnames(t.pred1) <- c("Predicted: No Vaccine", "Predicted: Vaccinated")
addmargins(A = t.pred1, FUN = list(Total=sum), quiet = TRUE)

# Confusion matrix for testing data set 
t.pred2 <-  table(prep_test$h1n1_vaccine, ypred2)
rownames(t.pred2) <- c("Actual:Not Vaccinated", "Actual: Vaccinated")
colnames(t.pred2) <- c("Predicted: Not Vaccinated", "Predicted: Vaccinated")
addmargins(A =t.pred2, FUN = list(Total=sum), quiet = TRUE)



```

# Naive Bayes Prediction 2
```{r}

pred_p2 <- predict(nb2, newdata=prep_train_rebal)
predicted.class_p2 <- factor(ifelse(pred_p > .5, "1","0"))
accuracy2 <- table(pred_p2, prep_train_rebal[,"h1n1_vaccine"])
sum(diag(accuracy1))/sum(accuracy2)
#printing the results
confusionMatrix(data=pred_p2, factor(prep_train_rebal$h1n1_vaccine), positive = '1')

#Evaluating the confusion matrix of the testing data set 
pred_p3 <- predict(nb2, newdata=prep_test)

#predicted.class_2 <- factor(ifelse(pred_p > 0.5, "1","0"))
accuracy3 <- table(pred_p3, prep_test[,"h1n1_vaccine"])
sum(diag(accuracy1))/sum(accuracy3)

#Printing the results
confusionMatrix(data=pred_p, factor(prep_test$h1n1_vaccine), positive = '1')



```


 Random Forest
```{r}
prep_train_rebal$h1n1_vaccine <- as.factor(prep_train_rebal$h1n1_vaccine)
prep_test$h1n1_vaccine <-as.factor(prep_test$h1n1_vaccine)
# had to convert the target variables to factors before running the code to run it as a classification model instead of regression.
library(randomForest)
h1n1_rf <- randomForest(h1n1_vaccine ~., data = prep_train_rebal, ntree = 100,proximity = TRUE)
print(h1n1_rf)
```
the out-of-bag error rate is 6.58%. Meaning that 93.42% of the data was predicted correctly.

```{r}
plot(h1n1_rf)
```
```{r}
plot(h1n1_rf$err.rate)
```


The Mean Decrease Gini below
```{r}
h1n1_rf$importance
```
what is mtry?

```{r}
h1n1_rf$mtry
```
why the accuracy? 

```{r}


p1<- predict(h1n1_rf, prep_train_rebal)
confusionMatrix(p1, prep_train_rebal$h1n1_vaccine, positive = '1')

```
prepare new model based on the gini coefficiency? what does it say?

```{r}
h1n1_meangini <- randomForest(h1n1_vaccine ~ doctor_recc_h1n1+
opinion_h1n1_risk + opinion_h1n1_vacc_effective +
age_numeric + opinion_h1n1_sick_from_vacc + education_numeric +
h1n1_concern + census_msa_numeric + household_adults + household_children + income_poverty_numeric + h1n1_knowledge +
health_worker + employment_numeric +race_numeric+
sex_numeric, data = prep_train_rebal, ntree = 100,proximity = TRUE)
```

gini model confusion matrix shows that the model accuracy decreased using the Mean Decrease Gini
```{r}
pred_gini<- predict(h1n1_meangini, prep_train_rebal)
confusionMatrix(pred_gini, prep_train_rebal$h1n1_vaccine, positive = '1')

```

Random Forest Model validation
```{r}
p2<- predict(h1n1_rf, prep_test)
confusionMatrix(p2, prep_test$h1n1_vaccine, positive = '1')
```
Gini Model Decrease Validation

```{r}
pred_gini_test<- predict(h1n1_meangini, prep_test)
confusionMatrix(pred_gini_test, prep_test$h1n1_vaccine, positive = '1')
```